---
output: html_document
---
## Date

### Start simple
```{r, results='asis'}
d1=date()
d1

class(d1)

```

### Date class
```{r, results='asis'}
d2 = Sys.Date()
d2

class(d2)
```

### Formatting dates
```
%d = day as number (0-31),
%a = abbreviated weekday,
%A = unabbreviated weekday,
%m = month (00-12),
%b = abbreviated month, 
%B = unabbrevidated month,
%y = 2 digit year, 
%Y = four digit year
```

```{r, results='asis'}
## datetime <- strptime(paste(data[,1],data[,2]), format = "%d/%m/%Y %H:%M:%S")
## strptime(as.character(tmpData), format="%m/%d/%Y %H:%M:%S")
## tmp <- as.date(as.character(tmpData),format="%m/%d/%Y %H:%M:%S")
## format(hehe,"%Y"))
format(d2,"%a %b %d")
```

### Creating dates
```{r, results='asis'}
Sys.setlocale("LC_TIME", "C")
x = c("1jan1960", "2jan1960", "31mar1960", "30jul1960");
z = as.Date(x, "%d%b%Y")
z
```

### Lubridate
```{r, results='asis'}
library(lubridate)
ymd("20140108")

mdy("08/04/2013")

dmy("03-04-2013")

ymd_hms("2011-08-03 10:15:03")
```

## Base Graphics

- ??Artist??s palette?? model
- Start with blank canvas and build up from there
- Start with plot function (or similar)
- Use annotation functions to add/modify (text, lines, points, axis)


## Simple Base Graphics
```{r, results='asis'}
# hist
library(datasets)
hist(airquality$Ozone) 

# Scatterplot
library(datasets)
with(airquality, plot(Wind, Ozone))

# boxplot
library(datasets)
airquality <- transform(airquality, Month = factor(Month))
boxplot(Ozone ~ Month, airquality, xlab = "Month", ylab = "Ozone (ppb)")

# summary(data) = returns min, 1st quartile, median, mean, 3rd quartile, max
summary(airquality)

```

### Some Important Base Graphics Parameters

Many base plotting functions share a set of parameters. Here are a few
key ones:

- `pch`: the plotting symbol (default is open circle)

- `lty`: the line type (default is solid line), can be dashed, dotted, etc.

- `lwd`: the line width, specified as an integer multiple

- `col`: the plotting color, specified as a number, string, or hex
  code; the `colors()` function gives you a vector of colors by name

- `xlab`: character string for the x-axis label

- `ylab`: character string for the y-axis label


### Some Important Base Graphics Parameters

The `par()` function is used to specify *global* graphics parameters
that affect all plots in an R session. These parameters can be
overridden when specified as arguments to specific plotting functions.

- `las`: the orientation of the axis labels on the plot
- `bg`: the background color
- `mar`: the margin size
- `oma`: the outer margin size (default is 0 for all sides)
- `mfrow`: number of plots per row, column (plots are filled row-wise) 
- `mfcol`: number of plots per row, column (plots are filled column-wise)


```{r, results='asis'}
par("lty")

par("col")

par("bg")
```


### Base Plotting Functions

- `plot`: make a scatterplot, or other type of plot depending on the
  class of the object being plotted

- `lines`: add lines to a plot, given a vector x values and a
  corresponding vector of y values (or a 2-column matrix); this
  function just connects the dots

- `points`: add points to a plot
- `text`: add text labels to a plot using specified x, y coordinates
- `title`: add annotations to x, y axis labels, title, subtitle, outer margin 
- `mtext`: add arbitrary text to the margins (inner or outer) of the plot 
- `axis`: adding axis ticks/labels


### Base Plot with Regression Line and Annotation

```{r, results='asis'}
with(airquality, plot(Wind, Ozone, main = "Ozone and Wind in New York City",  type = "n"))
with(subset(airquality, Month == 5), points(Wind, Ozone, col = "blue"))
with(subset(airquality, Month != 5), points(Wind, Ozone, col = "red"))
legend("topright", pch = 1, col = c("blue", "red"), legend = c("May", "Other Months"))

model <- lm(Ozone ~ Wind, subset(airquality, Month == 5))
abline(model, lwd = 2, col = "blue")

model2 <- lm(Ozone ~ Wind, subset(airquality, Month != 5))
abline(model2, lwd = 2, col = "red")
```


### multiple scatter plots
```{r, results='asis'}
par(mfrow = c(1, 2))
with(airquality, {
    plot(Wind, Ozone, main = "Ozone and Wind")
    plot(Solar.R, Ozone, main = "Ozone and Solar Radiation")
})
```

### Graphics Device
```{r, results='hide', echo=TRUE}
pdf(file = "myplot.png")  ## Open PDF device; create 'myplot.pdf' in my working directory
## Create plot and send to a file (no plot appears on screen)
with(faithful, plot(eruptions, waiting))
title(main = "Old Faithful Geyser data")  ## Annotate plot; still nothing on screen
dev.off()  ## Close the PDF file device
## Now you can view the file 'myplot.pdf' on your computer
```

## Plotting Lattice

- Plots are created with a single function call (xyplot, bwplot, etc.)
- Most useful for conditioning types of plots: Looking at how $y$ changes with $x$ across levels of $z$
- Things like margins/spacing set automatically because entire plot is specified at once
- Good for putting many many plots on a screen

### Lattice Functions

- `xyplot`: this is the main function for creating scatterplots 
- `bwplot`: box-and-whiskers plots (??boxplots??)
- `histogram`: histograms
- `stripplot`: like a boxplot but with actual points 
- `dotplot`: plot dots on "violin strings"
- `splom`: scatterplot matrix; like `pairs` in base plotting system 
- `levelplot`, `contourplot`: for plotting "image" data


### Lattice Functions

Lattice functions generally take a formula for their first argument, usually of the form

```r
xyplot(y ~ x | f * g, data)
```

- We use the *formula notation* here, hence the `~`.

- On the left of the ~ is the y-axis variable, on the right is the
  x-axis variable

- f and g are _conditioning variables_ ?? they are optional
  - the * indicates an interaction between two variables

- The second argument is the data frame or list from which the
  variables in the formula should be looked up

  - If no data frame or list is passed, then the parent frame is used.

- If no other arguments are passed, there are defaults that can be used.

```{r, results='asis'}
library(lattice)
library(datasets)
## Simple scatterplot
xyplot(Ozone ~ Wind, data = airquality)
```


### Simple Lattice Plot

```{r, results='asis'}
library(datasets)
library(lattice)
## Convert 'Month' to a factor variable
airquality <- transform(airquality, Month2 = factor(Month))
## Month is conditioning variables 
xyplot(Ozone ~ Wind | Month, data = airquality, layout = c(5, 1))
```

### Lattice Panel Function

```{r, results='asis'}
set.seed(10)
x <- rnorm(100)
f <- rep(0:1, each = 50)
y <- x + f - f * x + rnorm(100, sd = 0.5)
f <- factor(f, labels = c("Group 1", "Group 2"))

## Custom panel function
xyplot(y ~ x | f, panel = function(x, y, ...) {
    panel.xyplot(x, y, ...)  ## First call the default panel function for 'xyplot'
    panel.abline(h = median(y), lty = 2)  ## Add a horizontal line at the
    panel.lmline(x, y, col = 2)  ## Overlay a simple linear regression line
})
```

### Plotting Math Substituting Annotation

substitute(expr, env)

```{r, results='asis'}
x <- rnorm(100)
y <- x + rnorm(100, sd = 0.5)
plot(x, y,
     xlab=substitute(bar(x) == k, list(k=mean(x))),
     ylab=substitute(bar(y) == k, list(k=mean(y)))
     )

```

## ggplot2 Plotting System

- Split the difference between base and lattice
- Automatically deals with spacings, text, titles but also allows you to annotate by ??adding??
- Superficial similarity to lattice but generally easier/more intuitive to use
- Default mode makes many choices for you (but you can customize!)

### The Basics qplot()

#### Modifying aesthetics && Adding a geom && Histograms
```{r, results='asis'}
library(ggplot2)
# factor(mpg$drv)
qplot(displ, hwy, data = mpg, color = drv, geom = c("point", "smooth"), fill = drv)
```

#### Facets
```{r, results='asis'}
## horizental
## qplot(displ, hwy, data = mpg, facets = drv ~ .)

## portrait
qplot(displ, hwy, data = mpg, facets = . ~ drv)
```

#### Density Smootg
```{r, results='asis'}
## shape = drv
qplot(hwy, data = mpg, geom = "density", color = drv)
```


### Summary of qplot()

- The qplot() function is the analog to plot() but with many built-in features
- Syntax somewhere in between base/lattice
- Produces very nice graphics, essentially publication ready (if you like the design)
- Difficult to go against the grain/customize (don??t bother; use full ggplot2 power in that case)



### ggplot2

- data(package="ggplot2")
- data(packageName)

#### Basic Components

- A data frame
- aesthetic mappings: how data are mapped to color, size
- geoms: geometric objects like points, lines, shapes.
- facets: for conditional plots.
- stats: statistical transformations like binning, quantiles, smoothing.
- scales: what scale an aesthetic map uses (example: male = red, female = blue).
- coordinate system
- Plots are built up in layers
    - Plot the data
    - Overlay a summary
    - Metadata and annotation

#### Building Up in Layers
```{r, results='asis'}
g <- ggplot(mpg,aes(displ, hwy))
summary(g)
print(g)
g + geom_point()
```

#### Adding More Layers
```{r, results='asis'}
g + geom_point() + geom_smooth(method="lm")
g + geom_point() + facet_grid(. ~ drv) + geom_smooth(method = "lm")
```


#### Annotation

- Labels: xlab(), ylab(), labs(), ggtitle()
- Each of the ??geom?? functions has options to modify
- For things that only make sense globally, use theme()
- Example: theme(legend.position = "none")
- Two standard appearance themes are included
- theme_gray(): The default theme (gray background)
- theme_bw(): More stark/plain

#### Modifying Aesthetics
```{r, results='asis'}
## g + geom_point(aes(color = drv), size = 4, alpha = 1/2)
g + geom_point(color="steelblue", size = 4, alpha = 1/2)
```

```
## Modifying Labels
g + geom_point(aes(color = drv)) + labs(title = "my title") + 
+     labs(x = expression(" " * displ), y = "hwy")

## Customizing the Smooth
g + geom_point(aes(color = drv), size = 2, alpha = 1/2) + 
  geom_smooth(size = 4, linetype = 3, method = "lm", se = FALSE)
```
#### Changing the Theme
```
g + geom_point(aes(color = drv)) + theme_bw(base_family = "Times")
```

#### Axis Limit

```{r, results='asis'}
testdat <- data.frame(x = 1:100, y = rnorm(100))
testdat[50,2] <- 100  ## Outlier!
plot(testdat$x, testdat$y, type = "l", ylim = c(-3,3))

g <- ggplot(testdat, aes(x = x, y = y))
g + geom_line()
```


## Summary

- Base: "artist's palette" model

- Lattice: Entire plot specified by one function; conditioning

- ggplot2: Mixes elements of Base and Lattice
