---
title: "ReproducibleResearch"
author: "Anna"
date: "2016年4月9日"
output: html_document
---


* **`knitr` package**
	* written by YihuiXie, built into RStudio
	* support Markdown, LaTeX, HTML as documentation languages
	* Exports PDF/HTML
	* evaluates R markdown documents and return/records the results, and write out a Markdown files
	* Markdown file can then be converted into HTML using markdown package
	* solidify package converts R markdown into presentation slides
* In RStudio, create new R Markdown files by clicking New $\rightarrow$ R Markdown
	* `========` = indicates title of document (large text)
	* `$expression$` = indicates LaTeX expression/formatting
	* \``text`\` = changes text to code format (typewriter font)
	* \`\`\``{r name, echo = FALSE, results = hide}...`\`\`\` = R code chunk
		* `name` = name of the code chunk
		* `echo = FALSE` = turns off the echo of the R code chunk, which means display only the result
		* ***Note**: by default code in code chunk is echoed = print code AND results *
		* `results = hide` = hides the results from being placed in the markdown document
	* inline text computations
		* \``r` `variable`\` = prints the value of that variable directly inline with the text
	* incorporating graphics
		* \`\`\``{r scatterplot, fig.height = 4, fig.width = 6} ... plot() ... `\`\`\` = inserts a plot into markdown document
			* `scatterplot` = name of this code chunk (can be anything)
			* `fig.height = 4` = adjusts height of the figure, specifying this alone will produce a rectangular plot rather than a square one by default
			* `fig.width = 6` = adjusts width of the figure
		* knitr produces HTML, with the image embedded in HTML using base64 encoding
			* does not depend on external image files
			* not efficient but highly transportable
	* incorporating tables (xtable package: `install.packages("xtable")`)
		* `xtable`prints the table in html format, which is better presented than plain text normally

```{r showtable, results = "asis", fig.align = 'center'}
library(datasets)
## library(xtable)
fit <- lm(Ozone ~ Wind + Temp + Solar.R, data = airquality)
## xt <- xtable(summary(fit))
## print(xt, "html")
```

* **setting global options**
	* \`\`\``{r setoptions, echo = FALSE} opts_chunk$set(echo = FALSE, results = "hide")`\`\`\` = sets the default option to not print the code/results unless otherwise specified
* **common options**
	* **output**: `results = "asis"` OR `"hide"`
		* `"asis"` = output to stay in original format and not compiled into HTML
	* **output**: `echo = TRUE` OR `FALSE`
	* **figures**: `fig.height = numeric`
	* **figures**: `fig.width = numeric`
* **caching computations**
	* add argument to code chunk: `cache = TRUE`
	* computes and stores result of code the first time it is run, and calls the stored result directly from file for each subsequent call
	* useful for complex computations
	* caveats:
		* if data/code changes, you will need to re-run cached code chunks
		* dependencies not checked explicitly (changes in other parts of the code $\rightarrow$ need to re-run the cached code)
		* if code does something outside of the document (i.e. produce a png file), the operation cannot be cached
* **"Knit HTML"** button to process the document
	* alternatively, when not in RStudio, the process can be accomplished through the following

```r
library(knitr)
setwd(<working directory>)
knit2html("document.Rmd")
browseURL("document.html")
```
```{r fig.height=4, message=FALSE, results='asis', cache=TRUE}
##  highlight=TRUE,
x <- rnorm(100)
y <- x + rnorm(100, sd = 0.5)
par(mar = c(5,4,1,1), las=1)
plot(x, y, main="My Simulated Data")
```

* **`Use R Results**
```{r echo=FALSE, results="hide"}
time <- format(Sys.time(), "%a %b %d %X %Y")
rand <-rnorm(1)
```

The current time is `r time`. My favorite random number is `r rand`.


