---
title: "StatisticalInference"
author: "Anna"
date: "2016年5月14日"
output: html_document
---

### Conditional Probability
* let B = an event so that $P(B) > 0$
* **conditional probability** of an event A, given B is defined as the probability that BOTH A and B occurring divided by the probability of B occurring $$P(A\:|\:B) = \frac{P(A \:\cap\: B)}{P(B)}$$
* if A and B are ***independent***, then $$P(A\:|\:B) = \frac{P(A)P(B)}{P(B)} = P(A)$$
* ***example***
	* for die roll, $A = \{1\}$, $B = \{1, 3, 5\}$, then $$P(1~|~Odd) = P(A\:|\:B) = \frac{P(A \cap B)}{P(B)} = \frac{P(A)}{P(B)} = \frac{1/6}{3/6} = \frac{1}{3}$$

### Baye's Rule
* definition $$P(B\:|\:A) = \frac{P(A\:|\:B)P(B)}{P(A\:|\:B)P(B)+P(A\:|\:B^c)P(B^c)}$$ where $B^c$ = corresponding probability of event $B$, $P(B^c) = 1 - P(B)$

### Random Variables
* **random variable** = numeric outcome of experiment
* **discrete** (what you can count/categories) = assign probabilities to every number/value the variable can take
	* coin flip, rolling a die, web traffic in a day
* **continuous** (any number within a continuum) = assign probabilities to the range the variable can take
	* BMI index, intelligence quotients
	* ***Note**: limitations of precision in taking the measurements may imply that the values are discrete, but we in fact consider them continuous *
+ `rbinom()`, `rnorm()`, `rgamma()`, `rpois()`, `runif()` = functions to generate random variables from the binomial, normal, Gamma, Poisson, and uniform distributions

	
	
### Probability Mass Function (PMF)
* evaluates the probability that the **discrete random variable** takes on a specific value
	- measures the chance of a particular outcome happening
	* always $\ge$ 0 for every possible outcome
	* $\sum$ possible values that the variable can take = 1
* `dbinom(k, n, p)` = return the probability of getting `k` successes out of `n` trials, given probability of success is `p`



### Probability Density Function (PDF)
* evaluates the probability that the **continuous random variable** takes on a specific value 
	* always $\ge$ 0 everywhere
	* total area under curve must = 1

* `dnorm()`, `dgamma()`, `dpois()`, `dunif()` = return probability of a certain value from the normal, Gamma, Poisson, and uniform distributions



### Cumulative Distribution Function (CDF)
* CDF of a random variable $X$ = probability that the random variable is $\le$ value $x$
	* $F(x) = P(X \le x)$ = applies when $X$ is discrete/continuous
* PDF = derivative of CDF
	* integrate PDF $\rightarrow$ CDF
		* `integrate(function, lower=0, upper=1)` $\rightarrow$ can be used to evaluate integrals for a specified range
* `pbinom()`, `pnorm()`, `pgamma()`, `ppois()`, `punif()` = returns the cumulative probabilities from 0 up to a specified value from the binomial, normal, Gamma, Poisson, and uniform distributions


### Survival Function
* survival function of a random variable $X$ = probability the random variable $> x$, complement of CDF
    * $S(x) = P(X > x) = 1 - F(x)$, where $F(x) =$ CDF


### Quantile
* the $\alpha^{th}$ quantile of a distribution with distribution function F = point $x_{\alpha}$
	* $F(x_{\alpha}) = \alpha$
	* percentile = quantile with $\alpha$ expressed as a percent
	* median = 50^th^ percentile
	* $\alpha$% of the possible outcomes lie below it

* `qbeta(quantileInDecimals, 2, 1)` = returns quantiles for beta distribution
    * works for `qnorm()`, `qbinom()`, `qgamma()`, `qpois()`, etc.


### Example
```{r, echo=TRUE}
## What is the probability that 75% or fewer of calls get addressed?
## x = 1 ,y =2 tangle
## F(x) = P(X <= x) = 1/2 * Base * Height = x square
## 1 * 0.75 * 2 * 0.75 / 2
1.5 * 0.75/2

## a b??
pbeta(0.75, 2, 1)

pbeta(c(0.4, 0.5, 0.75), 2, 1)

## quantitles 25%
qbeta(0.25, 2, 1)


```

### Variance
* **variance** = measure of spread or dispersion, the expected squared distance of the variable from its mean (expressed in $X$'s units$^2$)
	- as we can see from above, higher variances $\rightarrow$ more spread, lower $\rightarrow$ smaller spread
	* $Var(X) = E[(X-\mu)^2] = E[X^2] - E[X]^2$
	* **standard deviation** $= \sqrt{Var(X)}$ $\rightarrow$ has same units as X
	

### Sample Variance
* the **sample variance** is defined as $$S^2 = \frac{\sum_{i=1} (X_i - \bar X)^2}{n-1}$$


### Of the mean of random sample
居然没有分清方差和样本方差
样本均值的均值与总体均值相等，样本均值的方差是总体方差的1/n

* **distribution for mean of random samples**
	* expected value of the **mean** of distribution of means = expected value of the sample mean = population mean
		* $E[\bar X]=\mu$
	* expected value of the variance of distribution of means
		* $Var(\bar X) = \sigma^2/n$

* **standard error (SE)**
	* the standard error of the mean is defined as $$SE_{mean} = \sigma/\sqrt{n}$$
	* this quantity is effectively the standard deviation of the distribution of a statistic (i.e. mean)
	* represents variability of means

* **estimate of standard error**
    * the sample standard error of the mean is defined as $$S/\sqrt{n}$$
    * estimates how variable sample means ($n$ size) from the population are

### Example -Standard Normal
* variance = 1
* means of **n** standard normals (sample) have standard deviation = $1/\sqrt{n}$

```{r message = F, warning = F}
# specify number of simulations with 10 as number of observations per sample
nosim <- 1000; n <-10
# estimated standard deviation of mean
sd(apply(matrix(rnorm(nosim * n), nosim), 1, mean))
# actual standard deviation of mean of standard normals
1 / sqrt(n)
```

* `rnorm()` = generate samples from the standard normal
* `matrix()` = puts all samples into a nosim by $n$ matrix, so that each row represents a simulation with `nosim` observations
* `apply()` = calculates the mean of the $n$ samples
* `sd()` = returns standard deviation


### Example - Standard Uniform
* standard uniform $\rightarrow$ triangle straight line distribution $\rightarrow$ mean = 1/2 and variance = 1/12
* means of random samples of $n$ uniforms have have standard deviation of $1/\sqrt{12 \times n}$


```{r message = F, warning = F}
# estimated standard deviation of the sample means
sd(apply(matrix(runif(nosim * n), nosim), 1, mean))
# actual standard deviation of the means
1/sqrt(12*n)
```


### Example - Poisson
* $Poisson(x^2)$ have variance of $x^2$
* means of random samples of $n~ Poisson(4)$ have standard deviation of $2/\sqrt{n}$

```{r message = F, warning = F}
# estimated standard deviation of the sample means
sd(apply(matrix(rpois(nosim * n, lambda=4), nosim), 1, mean))
# actual standard deviation of the means
2/sqrt(n)
```


### Example - Bernoulli
* for $p = 0.5$, the Bernoulli distribution has variance of 0.25
* means of random samples of $n$ coin flips have standard deviations of $1 / (2 \sqrt{n})$

```{r message = F, warning = F}
# estimated standard deviation of the sample means
sd(apply(matrix(sample(0 : 1, nosim * n, replace = TRUE), nosim), 1, mean))
# actual standard deviation of the means
1/(2*sqrt(n))
```


### Likelihood
概率 用于在已知一些参数的情况下，预测接下来的观测所得到的结果，而
似然性 则是用于在已知某些观测所得到的结果时，对有关事物的性质的参数进行估计。

Poisson interval

```{r, echo=TRUE}
x <- 5; t <- 94.32; lambda <- x / t
round(lambda + c(-1, 1) * qnorm(.975) * sqrt(lambda / t), 3)

poisson.test(x, T = 94.32)$conf


exp(confint(glm(x ~ 1 + offset(log(t)), family = poisson(link = log))))

```



### Confidence intervals

* Gosset's $t$ distribution

```{r, echo=TRUE}
g1 <- sleep$extra[1:10]
g2 <- sleep$extra[11:20]
difference <- g2 - g1
mn <- mean(difference)
s <- sd(difference)
n <- 10
mn + c(-1, 1) * qt(0.975, n - 1) * s/sqrt(n)

##  why is the same
t.test(difference)$conf.int
```